Evolution of precopulatory and postcopulatory inbreeding behaviour

### 2016-09-09 ms_pre_post.Rmd
Add last Results section, and switch to a new first paragraph of results explaining figures

@@

We first test our hypothesis that females' ability to bias fertilisation with respect to kinship causes selection for polyandry by comparing simulations in which tendency for polyandry, inbreeding mating strategy, and inbreeding fertilisation strategy can all evolve to simulations in which inbreeding fertilisation strategy is fixed to random mating with respect to kinship, as present in Figures 1 and 2, respectively. Figure 1 shows changes in mean allele values of tendency for polyandry ($P_{a}$; red lines and shading), inbreeding mating strategy ($M_{a}$; blue lines and shading), and inbreeding fertilisation strategy ($F_{a}$; black lines and shading) over 10000 generations as calculated across all individuals within a focal population over 40 replicate simulations with identical starting conditions. Solid lines represent the grand mean of replicate mean allele values, and shading around lines shows standard deviations around the mean. Panels A-H show all possible cost combinations for tendency for polyandry ($c_{P}$), inbreeding mating strategy ($c_{M}$), and inbreeding fertilisation strategy ($c_{F}$). Figure 2 shows an equivalent figure with changes in allele values of tendency for polyandry (red lines and shading) and inbreeding mating strategy (blue lines and shading) only. However, in Figure 2, a horizontal black line where mean allele value of inbreeding fertilisation strategy equals zero shows that these allele values have been fixed. Figure 2 has only four panels because costs of inbreeding fertilisation strategy is irrelevant when inbreeding fertilisation strategy alleles have no effect and mating is random (i.e., $F_{p}=0$).

@@

Figure 4 will show evolution of $M_{p}$ given pre-existing post-copulatory inbreeding avoidance, and Figure 5 will show evolution of $P_{p}$ and $F_{p}$ given pre-existing pre-copulatory inbreeding avoidance.

@@

|-----|---------------------------------------|-------------|---------------|--------------
| **A** | **Trait**                           | **Allele**  | **Genotype**  | **Phenotype**
|       | Tendency for polyandry              | $P_{a}$     | $P_{g}$       | $P_{p}$
|       | Inbreeding mating strategy          | $M_{a}$     | $M_{g}$       | $M_{p}$
|       | Inbreeding fertilisation strategy   | $F_{a}$     | $F_{g}$       | $F_{p}$
|       |                                     |             |               |
| **B** | **Description**                     | **Parameter**  | | **Defaut value(s)**
|       | Cost of tendency for polyandry                 | $c_{P}$        | | $0$, $0.02$
|       | Cost of inbreeding mating strategy             | $c_{M}$        | | $0$, $0.02$
|       | Cost of inbreeding fertilisation strategy      | $c_{F}$        | | $0$, $0.02$
|       | Focal female's number of offspring             | $n$            | | $8$
|       | Log-linear slope of inbreeding depression      | $\beta$        | | $3$
|       | Adult male immigrants per generation           | $\rho$         | | $5$
|       | Female carrying capacity                       | $K_{f}$        | | $100$
|       | Male carrying capacity                         | $K_{m}$        | | $100$
|       | Mutation rate of alleles                       | $\mu$          | | $0.001$
|       | Standard deviation of mutation effect size     | $\mu_{SD}$     | | $\sqrt{1/20}$
|       |                                                |                | |
| **C** | **Description**                                                   | | **Variable** |
|       | Coefficient of kinship                                            | | $k$          |
|       | Focal female's number of mates                                    | | $N_{males}$  |
|       | Female $i$'s perceived absolute mate quality of male $j$          | | $Q^{m}_{ij}$ |
|       | Female $i$'s perceived relative mate quality of male $j$          | | $q^{m}_{ij}$ |
|       | Female $i$'s perceived absolute fertilisation quality of male $j$ | | $Q^{f}_{ij}$ |
|       | Female $i$'s perceived relative fertilisation quality of male $j$ | | $q^{f}_{ij}$ |
|       | Viability of a focal female's offspring                           | | $\Psi_{off}$ |

>>>>

We present our results by showing mean allele value changes over generations. All Figures 1-5 show changes in mean allele values of tendency for polyandry ($P_{a}$; red lines and shading), inbreeding mating strategy ($M_{a}$; blue lines and shading), and inbreeding fertilisation strategy ($F_{a}$; black lines and shading) over 10000 generations as calculated across all individuals within a focal population over 40 replicate simulations with identical starting conditions. Solid lines represent the grand mean of replicate mean allele values, and shading around lines shows standard deviations around the mean. Panels within figures show different cost combinations for tendency for polyandry ($c_{P}$), inbreeding mating strategy ($c_{M}$), and inbreeding fertilisation strategy ($c_{F}$). Figures have fewer than eight panels if some costs are not relevant because allele values are fixed, and fixed allele values are shown as horizontal lines where line height is set to the value of alleles at fixation. Supporting Information S1 presents correlations of mean allele values across replicate simulations, which were generally weak; we found no evidence suggesting strong evolutionary divergence among replicates with identical starting conditions (e.g., as might be the case if populations stochastically evolved toward inbreeding avoidance through mate choice or fertilisation, but not both), meaning that inspecting grand means of replicate allele values accurately reflects expected evolutionary dynamics over generations.

<!--- Figure 2 shows an equivalent figure with changes in allele values of tendency for polyandry (red lines and shading) and inbreeding mating strategy (blue lines and shading) only. However, in Figure 2, a horizontal black line where mean allele value of inbreeding fertilisation strategy equals zero shows that these allele values have been fixed. Figure 2 has only four panels because costs of inbreeding fertilisation strategy is irrelevant when inbreeding fertilisation strategy alleles have no effect and mating is random (i.e., $F_{p}=0$). --->

@@

We first test our hypothesis that females' ability to bias fertilisation with respect to kinship causes selection for polyandry by comparing simulations in which tendency for polyandry, inbreeding mating strategy, and inbreeding fertilisation strategy can all evolve to simulations in which inbreeding fertilisation strategy is fixed to random mating with respect to kinship, as present in Figures 1 and 2, respectively.

@@

Figure 4 shows change in inbreeding mating strategy allele values over generations when tendency for polyandry and inbreeding fertilisation strategy alleles were fixed for adaptive values of inbreeding avoidance $P_{a}=1$ and $F_{a}=-5$. Despite that females in these simulations could already avoid inbreeding after mating occurred, inbreeding mating strategy evolved towards inbreeding avoidance given costs of $c_{M}=0$ (Figure 4A) and $c_{M}=0.02$ (Figure 4B). We therefore did not find evidence that the existence of adaptive inbreeding avoidance through fertilisation bias in any way precluded the evolution of adaptive inbreeding avoidance through mate choice.

Figure 5 shows change in tendency for polyandry and inbreeding fertilisation strategy allele values over generations when inbreeding mating strategy alleles were fixed for adaptive inbreeding avoidance $M_{a}=-5$. Existence of adaptive inbreeding avoidance through mate choice precluded evolution of polyandry for all combinations of costs (Figure 5), meaning that evolution of adpative polyandry to avoid inbreeding through inbreeding fertilisation strategy was never predicted in populations where females were already avoiding inbreeding through mate choice. Further, while inbreeding fertilisation strategy evolved toward inbreeding avoidance ($F_{a}<0$) in the absence of costs (Figure 5A,B), allele values of inbreeding fertilisation strategy were not as negative as when adaptive inbreeding avoidance through mate choice was not fixed at the outset (Figures 1A,B,C,D). We therefore found strong evidence that the existence of adaptive inbreeding avoidance through pre-copulatory mate choice precluded evolution of adaptive inbreeding avoidance through post-copulatory inbreeding avoidance through fertilisation bias.

<!--- Figure 4 will show evolution of $M_{p}$ given pre-existing post-copulatory inbreeding avoidance, and Figure 5 will show evolution of $P_{p}$ and $F_{p}$ given pre-existing pre-copulatory inbreeding avoidance. --->

@@

|-------|-------------------------------------|-------------|---------------|--------------
| **A** | **Trait**                           | **Allele**  | **Genotype**  | **Phenotype**
|       | Tendency for polyandry              | $P_{a}$     | $P_{g}$       | $P_{p}$
|       | Inbreeding mating strategy          | $M_{a}$     | $M_{g}$       | $M_{p}$
|       | Inbreeding fertilisation strategy   | $F_{a}$     | $F_{g}$       | $F_{p}$
|       |                                     |             |               |
| **B** | **Description**                     | **Parameter**  | | **Defaut value(s)**
|       | Cost of tendency for polyandry                 | $c_{P}$        | | $0$, $0.02$
|       | Cost of inbreeding mating strategy             | $c_{M}$        | | $0$, $0.02$
|       | Cost of inbreeding fertilisation strategy      | $c_{F}$        | | $0$, $0.02$
|       | Focal female's number of offspring             | $n$            | | $8$
|       | Log-linear slope of inbreeding depression      | $\beta$        | | $3$
|       | Adult male immigrants per generation           | $\rho$         | | $5$
|       | Female carrying capacity                       | $K_{f}$        | | $100$
|       | Male carrying capacity                         | $K_{m}$        | | $100$
|       | Mutation rate of alleles                       | $\mu$          | | $0.001$
|       | Standard deviation of mutation effect size     | $\mu_{SD}$     | | $\sqrt{1/20}$
|       |                                                |                | |
| **C** | **Description**                                                   | | **Variable** |
|       | Coefficient of kinship                                            | | $k$          |
|       | Focal female's number of mates                                    | | $N_{males}$  |
|       | Female $i$'s perceived absolute mate quality of male $j$          | | $Q^{m}_{ij}$ |
|       | Female $i$'s perceived relative mate quality of male $j$          | | $q^{m}_{ij}$ |
|       | Female $i$'s perceived absolute fertilisation quality of male $j$ | | $Q^{f}_{ij}$ |
|       | Female $i$'s perceived relative fertilisation quality of male $j$ | | $q^{f}_{ij}$ |
|       | Viability of a focal female's offspring                           | | $\Psi_{off}$ |
